---
title: "Training Data Distribution"
format: gfm
author: "Heidi Rodenhizer"
---

::: content-hidden
# Set Up

```{r}
#| include: false
library(googleCloudStorageR)
library(sf)
library(tidyverse)
library(ggnewscale)
library(biscale)
library(viridis)
library(patchwork)
```

```{r}
#| include: false
theme_set(theme_bw())
```

# Define Functions

```{r}
classify_counts = function(count_data, breaks) {
  new_count_data = case_when(
    count_data == breaks[1] ~ "1",
    count_data <= breaks[3] ~ "2",
    count_data <= breaks[4] ~ "3",
    count_data > breaks[4] ~ "4"
  )
}
```

# Import Data

## Domain

```{r}
#| output: false
circumpolar = st_read("./domain/circumpolar_domain.geojson") |>
  st_transform(6931)
```

## Subregions

Ecoregions from Dinerstein et al. 2017 will be used as our subregions.

```{r}
#| output: false
subregions = st_read("./domain/circumpolar_subregions.geojson") |>
  st_transform(6931) |>
  select(ecoregion = ECO_NAME)
```

## Training Data

```{r}
train_meta = gcs_get_object(
  "gs://abrupt_thaw/RTS_MODEL_V2/DATA/TRAINING_DATA/metadata.csv"
)
```

## Split Data

```{r}
splits = jsonlite::fromJSON("./domain/data/training_v1.0_splits_summary.json")

splits_df = map2_df(
  splits[["splits"]][1:3],
  names(splits[["splits"]][1:3]),
  ~ .x |>
    pluck("ecoregions") |>
    unlist() |>
    as_tibble() |>
    rename(ecoregion = 1) |>
    mutate(group = .y)
)

splits_yaml = yaml::read_yaml("./domain/data/training_v1.0_splits.yaml")
train_regions = map2_df(
  splits_yaml,
  names(splits_yaml),
  ~ .x |>
    as_tibble() |>
    rename(ecoregion = 1) |>
    mutate(group = .y)
) |>
  left_join(subregions, by = "ecoregion") |>
  st_as_sf()
```

## Countries Map

```{r}
world = rnaturalearth::ne_countries(scale = "large", returnclass = "sf")
world = world # |>
# st_transform(st_crs(circumpolar)) |>
# st_crop(
#   subregions |>
#     st_bbox() |>
#     st_as_sfc() |>
#     st_buffer(dist = 300000) |>
#     st_bbox()
# )
```

```{r}
crop_poly <- tibble(
  geometry = st_sfc(st_point(c(0, 90)), crs = 'EPSG:4326')
) %>%
  st_sf() %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  st_buffer(dist = 5250000) %>%
  smoothr::densify(n = 3) %>%
  st_transform(crs = 'EPSG:4326')

# arctic_circle <- tibble(lon = c(-180, 180), lat = c(66, 66)) %>%
#   st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
#   summarise(geometry = st_combine(geometry)) %>%
#   st_cast("LINESTRING") %>%
#   smoothr::densify(max_distance = 1)

world_north <- world %>%
  st_make_valid() |>
  st_intersection(crop_poly) %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  select(name_long, continent, region_un, subregion)

world_north$geom[4] <- st_buffer(world_north$geom[4], 0.25)

world_north <- world_north %>%
  st_union(by_feature = TRUE)

# world_north_buffer <- world_north %>%
#   st_buffer(50000)

long_lines_from_coords <- function(lon, lat_min, lat_max) {
  line <- expand_grid(lon = lon, lat = c(lat_min, lat_max)) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lon)

  return(line)
}

long_lines <- seq(-180, 180, by = 60) %>%
  map(
    ~ long_lines_from_coords(.x, lat_min = st_bbox(crop_poly)[2], lat_max = 90)
  ) %>%
  bind_rows()

lat_lines_from_coords <- function(lat, lon_min, lon_max) {
  line <- expand_grid(lon = c(lon_min, lon_max), lat = lat) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lat)

  return(line)
}

lat_lines <- seq(50, 90, by = 10) %>%
  map(~ lat_lines_from_coords(.x, lon_min = -180, lon_max = 180)) %>%
  bind_rows()
```

:::

# Check Training Polygon Counts by Subregion

```{r}
region_train_count = train_meta |>
  summarise(
    RTSCount = n(),
    .by = RegionName
  ) |>
  mutate(
    RTSPercent = RTSCount / sum(RTSCount)
  ) |>
  arrange(-RTSPercent)
region_train_count
```

```{r}
#| echo: false
ggplot(
  region_train_count,
  aes(x = RegionName, y = RTSCount)
) +
  geom_col() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
#| echo: false
ggplot(
  region_train_count,
  aes(x = RegionName, y = RTSPercent)
) +
  geom_col() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
small_clusters_percent = region_train_count |>
  filter(RTSPercent < 0.1) |>
  summarise(TotalPercentSmallClusters = sum(RTSPercent))
small_clusters_percent
```

# Map Training Data

## Convert metadata to sf

```{r}
train_points = train_meta %>%
  st_as_sf(coords = c("centroid_lon", "centroid_lat"), crs = 4326) %>%
  st_transform(crs = 6931) %>%
  bind_cols(st_coordinates(.)) |>
  mutate(
    RTS = as.integer(case_when(
      TrainClass == "positive" ~ 1,
      TrainClass == "negative" ~ 0
    )),
    NoRTS = as.integer(case_when(
      TrainClass == "positive" ~ 0,
      TrainClass == "negative" ~ 1
    ))
  )

rts_points = train_points |>
  filter(TrainClass == "positive")

neg_points = train_points |>
  filter(TrainClass == "negative")

# train_bboxes = train_points |>
#   st_buffer(dist = 4.77 * 256, endCapStyle = "SQUARE")
```

```{r}
# st_write(train_points, "./domain/train_points.geojson", delete_dsn = TRUE)
```

## Create Hex Grid
### Count Scaled Hex

```{r}
class_breaks = c(0, 1, 100, 200)

train_hex = train_points |>
  st_make_grid(
    cellsize = sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000, # get short side length of hexagon from area == 10000 km^2, and convert to m
    square = FALSE
  ) |>
  st_as_sf() |>
  rename(geometry = x) |>
  st_join(train_points, left = FALSE) |>
  summarize(
    Count = n(),
    RTSCount = sum(RTS),
    NoRTSCount = sum(NoRTS),
    .by = c(geometry)
  ) |>
  rowwise() |>
  mutate(
    bi_class = str_flatten(
      c(
        classify_counts(RTSCount, breaks = class_breaks),
        classify_counts(NoRTSCount, breaks = class_breaks)
      ),
      collapse = "-"
    )
  ) |>
  ungroup() |>
  mutate(
    Buffer = (sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000) /
      2 *
      (0.9 - sqrt(Count / max(Count)) * 0.9), # use this ratio to scale the hexagons by total count: hexagon short side length * percentile of total count
    geometry_scaled = st_buffer(geometry, dist = Buffer * -1) # geometry of scaled hexagons
  ) |>
  st_join(
    subregions |>
      select(-COLOR),
    largest = TRUE
  ) |>
  rename(ecoregion = ECO_NAME)
```

### Region Hex

```{r}
regions_hex = train_hex |>
  st_join(train_regions, largest = TRUE) |>
  summarise(
    geometry = st_union(geometry),
    .by = c(group)
  ) |>
  mutate(
    geometry_offset = st_buffer(geometry, dist = -5000),
    group = factor(
      case_when(
        str_detect(group, "test") ~ "Testing",
        str_detect(group, "train") ~ "Training",
        str_detect(group, "val") ~ "Validation"
      ),
      levels = c("Training", "Validation", "Testing")
    )
  ) |>
  filter(!is.na(group))
```

## Examples for Map Labels

```{r}
examples = train_hex |>
  filter(
    RTSCount == max(RTSCount) |
      NoRTSCount == max(NoRTSCount) |
      abs(RTSCount - NoRTSCount) / Count ==
        min(abs(RTSCount - NoRTSCount) / Count)
  ) |>
  mutate(
    geometry_centroid = st_centroid(geometry),
    nudge_x = c(10, 10, 10),
    nudge_y = c(10, 10, 10)
  ) |>
  st_set_geometry("geometry_centroid") |>
  select(Count, RTSCount, NoRTSCount, nudge_x, nudge_y)
examples = examples %>%
  bind_cols(
    st_coordinates(.$geometry_centroid)
  ) |>
  rename(x_start = X, y_start = Y) |>
  mutate(
    x_end = x_start + c(800000, 630000, 180000),
    y_end = y_start + c(450000, -350000, -1100000),
    label = paste0("Positive: ", RTSCount, "\nNegative: ", NoRTSCount),
    label_total = paste0("Total: ", Count)
  )
```

## Map

### Custom palette

```{r}
# column 1
reds = colorRampPalette(c("#FFFFFF", "#AE3A4E"))
reds_4 <- reds(4)
# row 1
blues = colorRampPalette(c("#FFFFFF", "#4885C1"))
blues_4 <- blues(4)
# purples = colorRampPalette(c("#FFFFFF", "#3F2949"))
# purples_4 <- purples(4)
# row 4
red_purples = colorRampPalette(c("#AE3A4E", "#3F2949"))
red_purples_4 = red_purples(4)
# column 4
blue_purples = colorRampPalette(c("#4885C1", "#3F2949"))
blue_purples_4 = blue_purples(4)
# column 2 from blues and red_purples
column2s = colorRampPalette(c(blues_4[2], red_purples_4[2]))
column2s_4 <- column2s(4)
# column 3 from blues and red_purples
column3s = colorRampPalette(c(blues_4[3], red_purples_4[3]))
column3s_4 <- column3s(4)


custom_pal4 <- c(
  # must be in this order! Otherwise it gives an error saying that the label formats are incorrect.
  "1-1" = "#FFFFFF", # 0 x, 0 y
  "2-1" = "#E4BDC4", # low x, 0 y
  "3-1" = "#C97B89", # mid x, 0 y
  "4-1" = "#AE3A4E", # high x, 0 y
  "1-2" = "#C2D6EA", # 0 x, low y
  "2-2" = "#AFA0B5", # low x, low y
  "3-2" = "#9C6A80", # mid x, low y
  "4-2" = "#89344C", # high x, low y (does this work?)
  "1-3" = "#85ADD5", # 0 x, mid y
  "2-3" = "#7A82A6", # low x, mid y
  "3-3" = "#6F5878", # mid x, mid y
  "4-3" = "#642E4A", # high x, mid y
  "1-4" = "#4885C1", # 0 x, high y
  "2-4" = "#456699", # low x, high y
  "3-4" = "#424771", # mid x, high y
  "4-4" = "#3F2949" # high x, high y
)
```

### Bivariate Legend

```{r}
total_counts = train_points |>
  st_drop_geometry() |>
  summarise(n = n(), .by = c(TrainClass)) |>
  mutate(
    n = paste("Total:", n),
    x = c(-1.35, 2),
    y = c(2, -0.9),
    angle = c(90, 0)
  )

bi_legend <- bi_legend(
  pal = custom_pal4,
  dim = 4,
  xlab = "RTS Positive (Count)",
  ylab = "RTS Negative (Count)",
  size = 5,
  breaks = list(
    "bi_x" = c(class_breaks, max(train_hex |> pull(RTSCount))),
    "bi_y" = c(class_breaks, max(train_hex |> pull(NoRTSCount)))
  )
) +
  geom_text(
    data = total_counts,
    aes(x = x, y = y, angle = angle, label = n),
    size = 1.85
  ) +
  coord_fixed(
    xlim = c(0.5, 3.5),
    ylim = c(0.5, 3.5),
    clip = "off"
  ) +
  theme(
    plot.margin = margin(6, 5, 6, 10),
    plot.background = element_rect(
      fill = fill_alpha("white", 0)
    )
  )
# bi_legend
```

### Size Legend

```{r}
hex_legend_data = train_hex |>
  slice(rep(1, 3))
bounds = st_bbox(hex_legend_data$geometry)
hex_legend_data = hex_legend_data |>
  mutate(
    geometry = geometry - c(bounds$xmin, bounds$ymin),
    Count = c(
      max(train_hex$Count),
      (max(train_hex$Count) - 1) * 0.5 + 1,
      1
    ),
    Buffer = (sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000) /
      2 *
      (0.9 - sqrt(Count / max(Count)) * 0.9), # use this ratio to scale the hexagons by total count: hexagon short side length * percentile of total count
    geometry_scaled = st_buffer(geometry, dist = Buffer * -1) # geometry of scaled hexagons
  ) |>
  rowwise() |>
  mutate(
    geometry_scaled = geometry_scaled - c(0, st_bbox(geometry_scaled)$ymin)
  )

# hex_legend_data$geometry_scaled[[1]] = hex_legend_data$geometry_scaled[[1]] -
#   c(0, st_bbox(hex_legend_data$geometry_scaled[[1]])$ymin)
arrow = tibble(
  x = 0 - 100000,
  y = 0 + 15000,
  xend = 0 - 100000,
  yend = st_bbox(hex_legend_data)["ymax"] - 100000
)

size_legend = ggplot() +
  geom_sf(
    data = hex_legend_data,
    aes(
      geometry = geometry_scaled,
      fill = Count
    ),
    color = "transparent"
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#3F2949"
  ) +
  geom_segment(
    data = arrow,
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend
    ),
    arrow = arrow(length = unit(3, "points"), angle = 45),
    linewidth = 0.2
  ) +
  geom_text(
    aes(
      x = st_bbox(hex_legend_data)["xmax"] -
        (st_bbox(hex_legend_data)["xmax"] - arrow$x) / 2,
      y = arrow$y - 100000,
      label = "Total Count"
    ),
    hjust = 0.5,
    size = 1.75,
    # angle = 90
  ) +
  coord_sf(clip = "off") +
  theme_void() +
  theme(
    legend.position = "none"
  )
size_legend
```

### Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
train_hexplot = ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.25
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.25
  ) +
  geom_sf(
    color = 'transparent',
    fill = 'gray90'
  ) +
  geom_sf(
    data = train_hex,
    aes(geometry = geometry),
    fill = "transparent",
    color = "gray95",
    linewidth = 0.5
  ) +
  geom_sf(
    data = train_hex,
    aes(geometry = geometry_scaled, fill = bi_class),
    color = "transparent",
    alpha = 1
  ) +
  bi_scale_fill(
    pal = custom_pal4,
    dim = 4,
    guide = "none"
  ) +
  geom_sf(
    data = regions_hex,
    aes(
      geometry = geometry_offset,
      linetype = group
    ),
    color = "black",
    fill = "transparent"
  ) +
  scale_linetype_manual(
    name = "Training\nGroup",
    # guide = guide_legend(reverse = TRUE),
    values = c("solid", "longdash", "dotted")
  ) +
  geom_segment(
    data = examples,
    aes(x = x_start, xend = x_end, y = y_start, yend = y_end)
  ) +
  geom_text(
    data = examples,
    aes(
      x = x_end + c(50000, 50000, -350000),
      y = y_end + c(0, 0, -150000),
      label = label
    ),
    size = 2,
    hjust = 0,
    lineheight = 1
  ) +
  geom_text(
    data = examples,
    aes(
      x = x_end + c(50000, 50000, -350000),
      y = y_end + c(rep(-250000, 2), -250000 - 150000),
      label = label_total
    ),
    size = 2,
    fontface = "bold",
    hjust = 0,
    lineheight = 1
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  # geom_sf(
  #   data = full_join(
  #     subregions,
  #     train_regions |> st_drop_geometry(),
  #     by = "ecoregion"
  #   ),
  #   aes(color = is.na(group)),
  #   fill = "transparent"
  # ) +
  # scale_color_manual(
  #   name = "Region not in\nsplits.yaml",
  #   values = c("black", "red")
  # ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void() +
  theme(
    legend.title = element_blank(),
    # legend.title = element_text(size = 5.5),
    legend.text = element_text(size = 5, margin = margin(0, 3, 0, 0)),
    legend.text.position = "left",
    legend.position = "inside",
    legend.position.inside = c(1, 0.16),
    legend.justification = c(1, 0),
    legend.key.size = unit(13, "points"),
    legend.key.spacing.x = unit(0, "points")
    # plot.margin = margin(10, 10, 10, 10) # trying to fix the outer circle getting cut off at top, bottom, left, and right
  )

bi_legend_location = c(
  left = 0.81,
  bottom = 0,
  right = 1,
  top = 0.145
)

size_legend_location = c(
  left = 0.75,
  bottom = 0.02,
  right = 0.8,
  top = 0.075
)

train_hexplot = train_hexplot +
  inset_element(
    bi_legend,
    left = bi_legend_location["left"],
    bottom = bi_legend_location["bottom"],
    right = bi_legend_location["right"],
    top = bi_legend_location["top"]
  ) +
  inset_element(
    size_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
train_hexplot
```

```{r}
ggsave(
  "./plots/training_data_map.png",
  train_hexplot,
  height = 6.5,
  width = 6.5
)
```