---
title: "Training Data Distribution"
format: gfm
author: "Heidi Rodenhizer"
---

::: content-hidden
# Set Up

```{r}
#| include: false
library(googleCloudStorageR)
library(sf)
library(tidyverse)
library(ggnewscale)
library(biscale)
library(viridis)
library(patchwork)
```

```{r}
#| include: false
theme_set(theme_bw())
```

# Define Functions

```{r}
classify_counts <- function(count_data, breaks) {
  new_count_data <- case_when(
    count_data == breaks[1] ~ "1",
    count_data <= breaks[3] ~ "2",
    count_data <= breaks[4] ~ "3",
    count_data > breaks[4] ~ "4"
  )
}
```

# Import Data

## Domain

```{r}
#| output: false
circumpolar <- st_read("./domain/circumpolar_domain.geojson") |>
  st_transform(6931)
```

## Training Data
It should be noted that the RegionName in metadata.csv does not always correspond to the correct subregion from Dinerstein et al. 2017, as some of the training tiles fell outside of the extent of the circumpolar_subregions.geojson file that was used to determine the RegionName. This does not present a problem for model training as the RegionName from metadata.csv was used in the creation of splits.yaml.
```{r}
# train_meta = gcs_get_object(
#   "gs://rts-mapping-v2/training/v1.0/metadata.csv"
# ) # This causes R to crash, but is the correct source file.
train_meta <- read.csv(
  "./domain/data/training_v1.0_metadata.csv"
) # read_csv causes R to crash
```

## Split Data

```{r}
# Splits from "gs://rts-mapping-v2/training/v1.0/splits.yaml"
splits <- yaml::read_yaml("./domain/data/training_v1.0_splits.yaml")
splits_df <- map2_df(
  splits[1:3],
  names(splits)[1:3],
  ~ .x |>
    as_tibble() |>
    rename(ecoregion = 1) |>
    mutate(group = .y)
)
```

## Countries Map

```{r}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
crop_poly <- tibble(
  geometry = st_sfc(st_point(c(0, 90)), crs = 'EPSG:4326')
) %>%
  st_sf() %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  st_buffer(dist = 5250000) %>%
  smoothr::densify(n = 3) %>%
  st_transform(crs = 'EPSG:4326')

world_north <- world %>%
  st_make_valid() |>
  st_intersection(crop_poly) %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  select(name_long, continent, region_un, subregion)

world_north$geom[4] <- st_buffer(world_north$geom[4], 0.25)

world_north <- world_north %>%
  st_union(by_feature = TRUE)

long_lines_from_coords <- function(lon, lat_min, lat_max) {
  line <- expand_grid(lon = lon, lat = c(lat_min, lat_max)) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lon)

  return(line)
}

long_lines <- seq(-180, 180, by = 60) %>%
  map(
    ~ long_lines_from_coords(.x, lat_min = st_bbox(crop_poly)[2], lat_max = 90)
  ) %>%
  bind_rows()

lat_lines_from_coords <- function(lat, lon_min, lon_max) {
  line <- expand_grid(lon = c(lon_min, lon_max), lat = lat) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lat)

  return(line)
}

lat_lines <- seq(50, 90, by = 10) %>%
  map(~ lat_lines_from_coords(.x, lon_min = -180, lon_max = 180)) %>%
  bind_rows()
```

:::

# Check Training Polygon Counts by Subregion

```{r}
region_train_count <- train_meta |>
  summarise(
    RTSCount = n(),
    .by = RegionName
  ) |>
  mutate(
    RTSPercent = RTSCount / sum(RTSCount)
  ) |>
  arrange(-RTSPercent)
region_train_count
```

```{r}
#| echo: false
ggplot(
  region_train_count,
  aes(x = RegionName, y = RTSCount)
) +
  geom_col() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
#| echo: false
ggplot(
  region_train_count,
  aes(x = RegionName, y = RTSPercent)
) +
  geom_col() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
small_clusters_percent <- region_train_count |>
  filter(RTSPercent < 0.1) |>
  summarise(TotalPercentSmallClusters = sum(RTSPercent))
small_clusters_percent
```

# Map Training Data

## Convert metadata to sf

```{r}
train_points <- train_meta %>%
  left_join(splits_df, by = c("RegionName" = "ecoregion")) %>%
  st_as_sf(coords = c("centroid_lon", "centroid_lat"), crs = 4326) %>%
  st_transform(crs = 6931) %>%
  bind_cols(st_coordinates(.)) |>
  mutate(
    RTS = as.integer(case_when(
      TrainClass == "positive" ~ 1,
      TrainClass == "negative" ~ 0
    )),
    NoRTS = as.integer(case_when(
      TrainClass == "positive" ~ 0,
      TrainClass == "negative" ~ 1
    ))
  ) |>
  st_as_sf()

rts_points <- train_points |>
  filter(TrainClass == "positive")

neg_points <- train_points |>
  filter(TrainClass == "negative")
```

```{r}
# st_write(train_points, "./domain/train_points.geojson", delete_dsn = TRUE)
```

## Create Hex Grid
### Count Scaled Hex

```{r}
class_breaks <- c(0, 1, 100, 200)

train_hex <- train_points |>
  st_make_grid(
    cellsize = sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000, # get short side length of hexagon from area == 10000 km^2, and convert to m
    square = FALSE
  ) |>
  st_as_sf() |>
  rename(geometry = x) |>
  st_join(train_points, left = FALSE) |>
  summarize(
    Count = n(),
    RTSCount = sum(RTS),
    NoRTSCount = sum(NoRTS),
    ecoregion = names(sort(table(RegionName), decreasing = TRUE)[1]),
    .by = c(geometry)
  ) |>
  rowwise() |>
  mutate(
    bi_class = str_flatten(
      c(
        classify_counts(RTSCount, breaks = class_breaks),
        classify_counts(NoRTSCount, breaks = class_breaks)
      ),
      collapse = "-"
    )
  ) |>
  ungroup() |>
  mutate(
    Buffer = (sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000) /
      2 *
      (0.9 - sqrt(Count / max(Count)) * 0.9), # use this ratio to scale the hexagons by total count: hexagon short side length * percentile of total count
    geometry_scaled = st_buffer(geometry, dist = Buffer * -1) # geometry of scaled hexagons
  )
```

### Region Hex

```{r}
regions_hex <- train_hex |>
  left_join(splits_df, by = c("ecoregion")) |>
  summarise(
    geometry = st_union(geometry),
    .by = c(group)
  ) |>
  mutate(
    geometry_offset = st_buffer(geometry, dist = -7000),
    group = factor(
      case_when(
        str_detect(group, "test") ~ "Testing",
        str_detect(group, "train") ~ "Training",
        str_detect(group, "val") ~ "Validation"
      ),
      levels = c("Training", "Validation", "Testing")
    )
  )
```

## Examples for Map Labels

```{r}
examples <- train_hex |>
  filter(
    RTSCount == max(RTSCount) |
      NoRTSCount == max(NoRTSCount) |
      RTSCount > 100 & NoRTSCount > 100
  ) |>
  mutate(
    geometry_centroid = st_centroid(geometry),
    nudge_x = c(10, 10, 10),
    nudge_y = c(10, 10, 10)
  ) |>
  st_set_geometry("geometry_centroid") |>
  select(Count, RTSCount, NoRTSCount, nudge_x, nudge_y)
examples <- examples %>%
  bind_cols(
    st_coordinates(.$geometry_centroid)
  ) |>
  rename(x_start = X, y_start = Y) |>
  mutate(
    x_end = x_start + c(630000, 630000, -1200000),
    y_end = y_start + c(-350000, -350000, 0),
    label = paste0("Positive: ", RTSCount, "\nNegative: ", NoRTSCount),
    label_total = paste0("Total: ", Count)
  )
```

## Map

### Custom palette

```{r}
# column 1
reds <- colorRampPalette(c("#FFFFFF", "#AE3A4E"))
reds_4 <- reds(4)
# row 1
blues <- colorRampPalette(c("#FFFFFF", "#4885C1"))
blues_4 <- blues(4)
# purples = colorRampPalette(c("#FFFFFF", "#3F2949"))
# purples_4 <- purples(4)
# row 4
red_purples <- colorRampPalette(c("#AE3A4E", "#3F2949"))
red_purples_4 <- red_purples(4)
# column 4
blue_purples <- colorRampPalette(c("#4885C1", "#3F2949"))
blue_purples_4 <- blue_purples(4)
# column 2 from blues and red_purples
column2s <- colorRampPalette(c(blues_4[2], red_purples_4[2]))
column2s_4 <- column2s(4)
# column 3 from blues and red_purples
column3s <- colorRampPalette(c(blues_4[3], red_purples_4[3]))
column3s_4 <- column3s(4)


custom_pal4 <- c(
  # must be in this order! Otherwise it gives an error saying that the label formats are incorrect.
  "1-1" = "#FFFFFF", # 0 x, 0 y
  "2-1" = "#E4BDC4", # low x, 0 y
  "3-1" = "#C97B89", # mid x, 0 y
  "4-1" = "#AE3A4E", # high x, 0 y
  "1-2" = "#C2D6EA", # 0 x, low y
  "2-2" = "#AFA0B5", # low x, low y
  "3-2" = "#9C6A80", # mid x, low y
  "4-2" = "#89344C", # high x, low y (does this work?)
  "1-3" = "#85ADD5", # 0 x, mid y
  "2-3" = "#7A82A6", # low x, mid y
  "3-3" = "#6F5878", # mid x, mid y
  "4-3" = "#642E4A", # high x, mid y
  "1-4" = "#4885C1", # 0 x, high y
  "2-4" = "#456699", # low x, high y
  "3-4" = "#424771", # mid x, high y
  "4-4" = "#3F2949" # high x, high y
)
```

### Bivariate Legend

```{r}
total_counts <- train_points |>
  st_drop_geometry() |>
  summarise(n = n(), .by = c(TrainClass)) |>
  mutate(
    n = paste("Total:", n),
    x = c(2, -1.35),
    y = c(-0.9, 2),
    angle = c(0, 90)
  )

bi_legend <- bi_legend(
  pal = custom_pal4,
  dim = 4,
  xlab = "RTS Positive (Count)",
  ylab = "RTS Negative (Count)",
  size = 5,
  breaks = list(
    "bi_x" = c(class_breaks, max(train_hex |> pull(RTSCount))),
    "bi_y" = c(class_breaks, max(train_hex |> pull(NoRTSCount)))
  )
) +
  geom_text(
    data = total_counts,
    aes(x = x, y = y, angle = angle, label = n),
    size = 1.85
  ) +
  coord_fixed(
    xlim = c(0.5, 3.5),
    ylim = c(0.5, 3.5),
    clip = "off"
  ) +
  theme(
    plot.margin = margin(6, 5, 6, 10),
    plot.background = element_rect(
      fill = fill_alpha("white", 0)
    )
  )
# bi_legend
```

### Size Legend

```{r}
hex_legend_data <- train_hex |>
  slice(rep(1, 3))
bounds <- st_bbox(hex_legend_data$geometry)
hex_legend_data <- hex_legend_data |>
  mutate(
    geometry = geometry - c(bounds$xmin, bounds$ymin),
    Count = c(
      max(train_hex$Count),
      (max(train_hex$Count) - 1) * 0.5 + 1,
      1
    ),
    Buffer = (sqrt((100000 * 2) / (3 * sqrt(3))) * sqrt(3) * 1000) /
      2 *
      (0.9 - sqrt(Count / max(Count)) * 0.9), # use this ratio to scale the hexagons by total count: hexagon short side length * percentile of total count
    geometry_scaled = st_buffer(geometry, dist = Buffer * -1) # geometry of scaled hexagons
  ) |>
  rowwise() |>
  mutate(
    geometry_scaled = geometry_scaled - c(0, st_bbox(geometry_scaled)$ymin)
  )

arrow <- tibble(
  x = 0 - 100000,
  y = 0 + 15000,
  xend = 0 - 100000,
  yend = st_bbox(hex_legend_data)["ymax"] - 100000
)

size_legend <- ggplot() +
  geom_sf(
    data = hex_legend_data,
    aes(
      geometry = geometry_scaled,
      fill = Count
    ),
    color = "transparent"
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#3F2949"
  ) +
  geom_segment(
    data = arrow,
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend
    ),
    arrow = arrow(length = unit(3, "points"), angle = 45),
    linewidth = 0.2
  ) +
  geom_text(
    aes(
      x = st_bbox(hex_legend_data)["xmax"] -
        (st_bbox(hex_legend_data)["xmax"] - arrow$x) / 2,
      y = arrow$y - 100000,
      label = "Total Count"
    ),
    hjust = 0.5,
    size = 1.75,
    # angle = 90
  ) +
  coord_sf(clip = "off") +
  theme_void() +
  theme(
    legend.position = "none"
  )
# size_legend
```

### Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
train_hexplot <- ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.25
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.25
  ) +
  geom_sf(
    color = 'transparent',
    fill = 'gray90'
  ) +
  geom_sf(
    data = train_hex,
    aes(geometry = geometry_scaled, fill = bi_class),
    color = "transparent",
    alpha = 1
  ) +
  geom_sf(
    data = train_hex,
    aes(geometry = geometry),
    fill = "transparent",
    color = "gray95",
    linewidth = 0.5
  ) +
  bi_scale_fill(
    pal = custom_pal4,
    dim = 4,
    guide = "none"
  ) +
  geom_sf(
    data = regions_hex,
    aes(
      geometry = geometry_offset,
      linetype = group
    ),
    color = "black",
    fill = "transparent"
  ) +
  scale_linetype_manual(
    name = "Training\nGroup",
    values = c("solid", "longdash", "dotted")
  ) +
  geom_segment(
    data = examples,
    aes(x = x_start, xend = x_end, y = y_start, yend = y_end),
    linewidth = 0.5
  ) +
  geom_text(
    data = examples,
    aes(
      x = x_end + c(50000, 50000, -900000),
      y = y_end + c(50000, 50000, 50000),
      label = label
    ),
    size = 2,
    hjust = 0,
    lineheight = 1
  ) +
  geom_text(
    data = examples,
    aes(
      x = x_end + c(50000, 50000, -900000),
      y = y_end + c(-200000, -200000, -200000),
      label = label_total
    ),
    size = 2,
    fontface = "bold",
    hjust = 0,
    lineheight = 1
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 5, margin = margin(0, 3, 0, 0)),
    legend.text.position = "left",
    legend.position = "inside",
    legend.position.inside = c(1, 0.17),
    legend.justification = c(1, 0),
    legend.key.size = unit(13, "points"),
    legend.key.spacing.x = unit(0, "points")
  )

bi_legend_location <- c(
  left = 0.81,
  bottom = 0,
  right = 1,
  top = 0.145
)

size_legend_location <- c(
  left = 0.75,
  bottom = 0.02,
  right = 0.8,
  top = 0.075
)

train_hexplot <- train_hexplot +
  inset_element(
    bi_legend,
    left = bi_legend_location["left"],
    bottom = bi_legend_location["bottom"],
    right = bi_legend_location["right"],
    top = bi_legend_location["top"]
  ) +
  inset_element(
    size_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
train_hexplot
```

```{r}
ggsave(
  "./plots/training_data_map.png",
  train_hexplot,
  height = 6.5,
  width = 6.5
)
```