---
title: "Training Data Distribution"
format: gfm
author: "Heidi Rodenhizer"
---

::: content-hidden
# Set Up

## Packages

```{r}
#| include: false
library(googleCloudStorageR)
library(sf)
library(terra)
library(tidyterra)
# library(arcgis)
library(tidyverse)
library(ggnewscale)
# library(biscale)
library(viridis)
library(patchwork)
```

```{r}
# token <- auth_code()
```

```{r}
#| include: false
theme_set(theme_bw())
```

## Colors

```{r}
reds <- colorRampPalette(c("#FFFFFF", "#AE3A4E"))
reds_10 <- reds(10)
red_min <- reds_10[2]
red_max <- reds_10[10]
yellow_reds <- colorRampPalette(c("#fff7aa", "#ff6600", "#800000"))
rts_colors <- colorRampPalette(c("#FFFFFF", "#BAB08C", "#9F6155", "#4D0001")) # From Greg's RTS color ramp in his style file
rts_palettes <- list(
  rts_colors_3 = rts_colors(3),
  rts_colors_4 = rts_colors(4),
  rts_colors_5 = rts_colors(5),
  rts_colors_6 = rts_colors(6),
  rts_colors_7 = rts_colors(7),
  rts_colors_8 = rts_colors(8),
  rts_colors_9 = rts_colors(9),
  rts_colors_10 = rts_colors(10)
)
yellow_red_palettes <- list(
  yellow_red_3 = yellow_reds(3),
  yellow_red_4 = yellow_reds(4),
  yellow_red_5 = yellow_reds(5),
  yellow_red_6 = yellow_reds(6),
  yellow_red_7 = yellow_reds(7),
  yellow_red_8 = yellow_reds(8),
  yellow_red_9 = yellow_reds(9),
  yellow_red_10 = yellow_reds(10)
)
# "#FFF7AA"
# "#FF9638"
# "#D44400"
# "#800000"
# write_csv(
#   tibble(
#     yellow_red_3 = c(yellow_reds(3), rep(NA, 7)),
#     yellow_red_4 = c(yellow_reds(4), rep(NA, 6)),
#     yellow_red_5 = c(yellow_reds(5), rep(NA, 5)),
#     yellow_red_6 = c(yellow_reds(6), rep(NA, 4)),
#     yellow_red_7 = c(yellow_reds(7), rep(NA, 3)),
#     yellow_red_8 = c(yellow_reds(8), rep(NA, 2)),
#     yellow_red_9 = c(yellow_reds(9), rep(NA, 1)),
#     yellow_red_10 = yellow_reds(10)
#   ),
#   "./visualization/data/color_ramps.csv"
# )
# write_csv(
#   tibble(
#     rts_colors_3 = c(rts_colors(3), rep(NA, 7)),
#     rts_colors_4 = c(rts_colors(4), rep(NA, 6)),
#     rts_colors_5 = c(rts_colors(5), rep(NA, 5)),
#     rts_colors_6 = c(rts_colors(6), rep(NA, 4)),
#     rts_colors_7 = c(rts_colors(7), rep(NA, 3)),
#     rts_colors_8 = c(rts_colors(8), rep(NA, 2)),
#     rts_colors_9 = c(rts_colors(9), rep(NA, 1)),
#     rts_colors_10 = rts_colors(10)
#   ),
#   "./visualization/data/rts_color_ramps.csv"
# )
```


# Define Functions

## scientific_10

```{r}
scientific_10 <- function(x) {
  parse(
    text = paste(
      "'\u2264'*",
      str_replace(
        scales::scientific_format()(x),
        "e(\\+?)(0?)",
        " %*% 10^"
      )
    )
  )
}
```

## classify_counts

```{r}
classify_counts <- function(count_data, breaks) {
  new_count_data <- case_when(
    count_data <= breaks[1] ~ 1,
    count_data <= breaks[2] ~ 2,
    count_data <= breaks[3] ~ 3,
    count_data <= breaks[4] ~ 4
  )
}
```

## calculate_buffer

```{r}
calculate_buffer <- function(column, hex_size, min_area = 0.05) {
  (
    sqrt((hex_size * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
  ) -
    (
      sqrt(
        (hex_size *
          ((column - 1) /
            (max(column) - 1) *
            (1 - min_area) +
            min_area) * #
          2) /
          sqrt(3)
      ) /
        2 *
        1000 # radius of a circle of size (10,000 km^2 * percentile of variable of interest) in m
    )
}
```

## create_hex_legend

```{r}
create_hex_legend <- function(
  dataset,
  hex_size,
  variable_name,
  n,
  breaks,
  colors
) {
  hex_legend_data <- dataset |>
    select(geometry) |>
    slice(rep(1, n))
  bounds <- st_bbox(hex_legend_data$geometry)
  hex_legend_data <- hex_legend_data |>
    mutate(
      geometry = geometry - c(bounds$xmin, bounds$ymin),
      count = seq(n, 1),
      geometry_scaled = st_buffer(
        geometry,
        dist = calculate_buffer(count, hex_size) * -1
      ) # geometry of scaled hexagons
    ) |>
    rowwise() |>
    mutate(
      geometry_scaled = geometry_scaled - c(0, st_bbox(geometry_scaled)$ymin)
    ) |>
    select(count, geometry, geometry_scaled)

  hex_legend_bbox <- st_bbox(hex_legend_data)

  # arrow <- tibble(
  #   x = 0 - 100000,
  #   y = 0 + 15000,
  #   xend = 0 - 100000,
  #   yend = st_bbox(hex_legend_data)["ymax"] - 100000
  # )

  # Format breaks to use scientific notation if they are large
  if (all(breaks >= 1000)) {
    breaks <- scientific_10(breaks)
  } else if (all(breaks <= 1)) {
    breaks <- paste("\u2264", format(round(breaks, 2), nsmall = 2))
  } else {
    breaks <- paste("\u2264", round(breaks))
  }

  labels <- tibble(
    x = rep(-20000, 4),
    xend = c(
      hex_legend_bbox["xmax"] -
        (hex_legend_bbox["xmax"] - hex_legend_bbox["xmin"]) /
          2,
      NA,
      NA,
      hex_legend_bbox["xmax"] -
        (hex_legend_bbox["xmax"] - hex_legend_bbox["xmin"]) /
          2
    ),
    y = seq(
      st_bbox(hex_legend_data$geometry_scaled[[4]])["ymax"],
      st_bbox(hex_legend_data)["ymax"],
      length.out = 4
    ),
    label = breaks
  )

  size_legend <- ggplot() +
    geom_sf(
      data = hex_legend_data,
      aes(
        geometry = geometry_scaled,
        fill = factor(count)
      ),
      color = "gray20",
      # color = "gray10"
    ) +
    scale_fill_manual(
      values = colors
    ) +
    # geom_segment(
    #   data = arrow,
    #   aes(
    #     x = x,
    #     y = y,
    #     xend = xend,
    #     yend = yend
    #   ),
    #   arrow = arrow(length = unit(3, "points"), angle = 45),
    #   linewidth = 0.2
    # ) +
    geom_segment(
      data = labels,
      aes(
        x = x,
        y = y,
        xend = xend,
        yend = y
      ),
      linewidth = 0.2
    ) +
    geom_text(
      data = labels,
      aes(
        x = x,
        y = y,
        label = label
      ),
      hjust = 1,
      nudge_x = -20000,
      size = 6
    ) +
    geom_text(
      aes(
        x = hex_legend_bbox["xmax"] +
          (hex_legend_bbox["xmax"] - hex_legend_bbox["xmin"]) * 0.2,
        y = hex_legend_bbox["ymin"] +
          (hex_legend_bbox["ymax"] - hex_legend_bbox["ymin"]) / 2,
        label = variable_name
      ),
      hjust = 0,
      size = 12
    ) +
    coord_sf(clip = "off") +
    theme_void() +
    theme(
      plot.margin = margin(0.1, 2.4, 0.1, 1.1, unit = "in"),
      legend.position = "none"
    )

  return(size_legend)
}
```

## plot_hex_grid

```{r}
plot_hex_grid <- function(
  hex_data,
  geometry_col,
  geometry_col_scaled,
  fill_col,
  colors
) {
  ggplot(world_north) +
    # Background (lon, lat, land)
    geom_sf(
      data = long_lines,
      color = 'gray85',
      linewidth = 0.1
    ) +
    geom_sf(
      data = lat_lines,
      color = 'gray85',
      linewidth = 0.1
    ) +
    geom_sf(
      color = 'gray85',
      fill = 'gray85'
    ) +

    # # Permafrost
    geom_spatraster(
      data = permafrost,
      aes(fill = NIEER_Probability),
      # maxcell = 10e+06,
      # alpha = 0.75
    ) +
    scale_fill_gradient(
      low = "#d8ebfd",
      high = "#88BBEE",
      na.value = "transparent"
    ) +
    new_scale_fill() +

    # Cities
    geom_sf(
      data = cities,
      shape = 21,
      color = "black",
      fill = "white",
      size = 2
    ) +
    geom_sf(
      data = cities,
      color = "black",
      size = 0.5
    ) +
    geom_sf_text(
      data = cities,
      aes(label = CITY_NAME),
      hjust = 1,
      position = position_nudge(x = -150000),
      color = "black",
      size = 2
    ) +

    # hex borders
    geom_sf(
      color = 'white',
      fill = 'transparent',
      linewidth = 0.1
    ) +

    # RTS Hex
    geom_sf(
      data = hex_data,
      aes(geometry = {{ geometry_col_scaled }}, fill = factor({{ fill_col }})),
      color = "transparent",
      # color = "gray10",
      alpha = 0.8
    ) +
    scale_fill_manual(
      name = "RTS Area",
      values = colors
    ) +

    # RTS Hex Outlines
    geom_sf(
      data = hex_data,
      aes(geometry = {{ geometry_col }}),
      fill = "transparent",
      color = "gray95",
      linewidth = 0.3
    ) +

    # Inference Region
    geom_sf(
      data = region_hex,
      color = "black",
      fill = "transparent",
      linewidth = 0.3
    ) +

    # Manual Circular Plot Border
    geom_sf(
      data = crop_poly,
      color = 'black',
      fill = 'transparent',
      linewidth = 0.25
    ) +

    # Aesthetic Options
    scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
    scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
    theme_void() +
    theme(
      legend.position = "none"
    )
}
```


# Import Data

## Domain

```{r}
#| output: false
circumpolar <- st_read("./domain/circumpolar_domain.geojson") |>
  st_transform(6931)
circumpolar_south <- st_read("./domain/circumpolar_south_domain.geojson") |>
  st_transform(6931)
```

## Predictions

```{r}
# gs://rts-mapping-v2-usw1/inference/2025q3_south/products/south_rts_candidates.gpkg
rts <- st_read("../RTSMapping_v2_products/2025q3/south_rts_candidates.gpkg") |>
  st_transform(6931)
```

## Hex Grid

```{r}
hex_grid <- st_read("./visualization/data/training_hex_grid.geojson")
```

## Planet Grids

```{r}
planet_grids <- st_read(
  "../circumpolar_planet_basemaps/data/circumpolar_south_planet_basemap_grids_2025.geojson"
) |>
  st_transform(6931)
```

## ARTS

```{r}
arts <- st_read(
  "../ARTS/ARTS_main_dataset/v.6.0.0/ARTS_main_dataset_v.6.0.0.geojson"
)
```

## Countries Map

```{r}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
crop_poly <- tibble(
  geometry = st_sfc(st_point(c(0, 90)), crs = 'EPSG:4326')
) %>%
  st_sf() %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  st_buffer(dist = 5250000) %>%
  smoothr::densify(n = 3) %>%
  st_transform(crs = 'EPSG:4326')

crop_bbox <- crop_poly |>
  st_transform(6931) |>
  st_bbox()

world_north <- world %>%
  st_make_valid() |>
  st_intersection(crop_poly) %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  select(name_long, continent, region_un, subregion)

world_north$geom[4] <- st_buffer(world_north$geom[4], 0.25)

world_north <- world_north %>%
  st_union(by_feature = TRUE)

long_lines_from_coords <- function(lon, lat_min, lat_max) {
  line <- expand_grid(lon = lon, lat = c(lat_min, lat_max)) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lon)

  return(line)
}

long_lines <- seq(-180, 180, by = 60) %>%
  map(
    ~ long_lines_from_coords(.x, lat_min = st_bbox(crop_poly)[2], lat_max = 90)
  ) %>%
  bind_rows()

lat_lines_from_coords <- function(lat, lon_min, lon_max) {
  line <- expand_grid(lon = c(lon_min, lon_max), lat = lat) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lat)

  return(line)
}

lat_lines <- seq(50, 90, by = 10) %>%
  map(~ lat_lines_from_coords(.x, lon_min = -180, lon_max = 180)) %>%
  bind_rows()

world_ocean <- crop_bbox |>
  st_as_sfc() |>
  st_buffer(dist = 150000, joinStyle = "MITRE", mitreLimit = 150000) |>
  st_transform(st_crs(world_north)) |>
  st_difference(
    world_north |>
      st_union()
  )

# st_write(
#   world_ocean,
#   "./visualization/data/world_oceans.geojson"
# )
```

## Cities

```{r}
cities <- st_read("../../../external_datasets/cities/cities.shp") |>
  st_transform(6931) |>
  filter(
    CITY_NAME %in%
      c("Anchorage", "Moscow", "Reykjavik", "Tromso", "Yakutsk", "Yellowknife")
  )
```

```{r}
# st_write(
#   cities,
#   "./visualization/data/arctic_cities.geojson"
# )
```

## Esri Multi-direction Hillshade

```{r}
# hillshd_srvr <- arc_open(
#   "https://elevation.arcgis.com/arcgis/rest/services/WorldElevation/Terrain/ImageServer",
#   token = token
# )

# hillshd <- arc_raster(
#   hillshd_srvr,
#   xmin = -180,
#   xmax = 180,
#   ymin = 43.9,
#   ymax = 90,
#   crs = 4326,
#   token = token
# )

# # hillshd = hillshd |>
# #   crop(ext(vect(crop_poly)))
```

## Permafrost

```{r}
crop_poly_vect <- crop_poly |>
  st_transform(6931) |>
  vect()

permafrost <- rast(
  "../rts_rep_analysis/input_data/NIEERPermafrost/NIEER_permafrost_dataset_released/NIEER_Probability.tif"
) |>
  aggregate(fact = 12) |>
  project("epsg:6931") |>
  crop(crop_poly_vect) |>
  mask(crop_poly_vect)
```

:::

# Data Prep

## RTS Centroids

```{r}
rts_pts <- rts |>
  st_centroid()

rts_pts_high <- rts_pts |>
  filter(conf_class == "high")
```


```{r}
# st_write(
#   rts_pts,
#   "./visualization/data/rts_centroids_2025.geojson"
# )

# st_write(
#   rts_pts_high,
#   "./visualization/data/rts_centroids_high_conf_2025.geojson"
# )
```

```{r}
ggplot(data = rts_pts_high) +
  geom_histogram(
    data = rts_pts,
    aes(
      x = area_m2,
      color = "All Confidence Levels",
      fill = "All Confidence Levels"
    )
  ) +
  geom_histogram(
    aes(
      x = area_m2,
      color = "High Confidence",
      fill = "High Confidence"
    )
  ) +
  scale_x_continuous(trans = "log10")

ggplot(data = rts_pts_high) +
  geom_histogram(
    data = rts_pts,
    aes(
      x = mean_prob,
      color = "All Confidence Levels",
      fill = "All Confidence Levels"
    )
  ) +
  geom_histogram(
    aes(
      x = mean_prob,
      color = "High Confidence",
      fill = "High Confidence"
    )
  )

ggplot(data = rts_pts_high) +
  geom_histogram(
    data = rts_pts,
    aes(
      x = max_prob,
      color = "All Confidence Levels",
      fill = "All Confidence Levels"
    )
  ) +
  geom_histogram(
    aes(
      x = max_prob,
      color = "High Confidence",
      fill = "High Confidence"
    )
  )

ggplot(data = rts_pts_high) +
  geom_point(
    aes(
      x = mean_prob,
      y = max_prob
    ),
    alpha = 0.2
  )
```

## AK RTS

```{r}
us <- world_north |>
  filter(name_long == "United States")
us_buffer <- us |>
  st_buffer(dist = 2 * 10^5)
alaska <- us |>
  st_crop(st_sfc(
    st_polygon(us_buffer$geometry[[1]][[1]]),
    crs = st_crs(world_north)
  ))

ak_rts_pred <- rts_pts |>
  st_join(alaska, join = st_within, left = FALSE) |>
  mutate(
    conf_class = factor(
      conf_class,
      levels = c("low", "medium", "high")
    )
  )

ak_rts_arts <- arts |>
  filter(TrainClass == "Positive") |>
  st_centroid() |>
  st_transform(6931) |>
  st_join(alaska, join = st_within, left = FALSE)

ggplot(alaska) +
  geom_sf() +
  geom_sf(
    data = ak_rts_arts,
    aes(color = "ARTS Observations"),
    size = 0.1
  ) +
  scale_color_manual(
    name = "ARTS\nObservations",
    values = c("red"),
    labels = NULL
  ) +
  new_scale_color() +
  geom_sf(
    data = ak_rts_pred,
    aes(color = conf_class),
    size = 0.1
  ) +
  scale_color_manual(
    name = "Model\nConfidence",
    values = c("gray80", "gray50", "black")
  ) +
  coord_sf(crs = 6396) # AK State Plane Zone 3

# ggsave(
#   "./plots/ak_rts_pts.png",
#   height = 2.7,
#   width = 6.5
# )

ak_rts_pred |>
  st_drop_geometry() |>
  summarise(
    count = n(),
    .by = conf_class
  ) |>
  bind_rows(
    tibble(
      conf_class = c("ARTS observations"),
      count = nrow(ak_rts_arts)
    )
  )
```

## RTS Hex Grid

```{r}
rts_hex <- hex_grid |>
  st_join(rts_pts_high, left = FALSE) |>
  summarize(
    total_count = n(),
    total_area = sum(area_m2),
    median_area = median(area_m2),
    mean_prob = sum(mean_prob * area_m2) / total_area,
    max_prob = sum(max_prob * area_m2) / total_area,
    .by = c(geometry)
  )

test_small_rts_hex <- hex_grid |>
  st_make_grid(
    cellsize = sqrt((50000 * 2) / (sqrt(3))) * 1000, # get short side length of hexagon from area == 100000 km^2, and convert to m
    square = FALSE
  ) |>
  st_as_sf() |>
  rename(geometry = x) |>
  st_join(rts_pts_high, left = FALSE) |>
  summarize(
    total_count = n(),
    total_area = sum(area_m2),
    median_area = median(area_m2),
    mean_prob = sum(mean_prob * area_m2) / total_area,
    max_prob = sum(max_prob * area_m2) / total_area,
    .by = c(geometry)
  )

ggplot() +
  geom_sf(
    data = test_small_rts_hex,
    aes(fill = total_count)
  ) +
  geom_sf(
    data = rts_hex,
    fill = "transparent"
  ) +
  scale_fill_viridis(trans = "log10")
```

Take a look at distribution of counts and areas - consider a log scale for color ramps on these.

```{r}
ggplot(data = rts_hex) +
  geom_histogram(aes(x = total_count)) +
  scale_x_continuous(trans = "log10")

ggplot(data = rts_hex) +
  geom_histogram(aes(x = total_area)) +
  scale_x_continuous(trans = "log10")

ggplot(data = rts_hex) +
  geom_histogram(aes(x = median_area)) +
  scale_x_continuous(trans = "log10")

ggplot(data = rts_hex) +
  geom_histogram(aes(x = mean_prob))

ggplot(data = rts_hex) +
  geom_histogram(aes(x = max_prob))
```

```{r}
total_count_breaks <- rgeoda::natural_breaks(4, rts_hex["total_count"])
total_count_breaks <- c(
  30,
  plyr::round_any(total_count_breaks[1:2], 50),
  total_count_breaks[3]
)

total_area_breaks <- rgeoda::natural_breaks(4, rts_hex["total_area"])
total_area_breaks <- c(
  200000,
  plyr::round_any(total_area_breaks[1:2], 500000),
  total_area_breaks[3]
)

median_area_breaks <- rgeoda::natural_breaks(4, rts_hex["median_area"])
median_area_breaks <- c(
  plyr::round_any(median_area_breaks[1:3], 5000),
  max(rts_hex$median_area)
)

mean_prob_breaks <- rgeoda::natural_breaks(4, rts_hex["mean_prob"])
mean_prob_breaks <- c(
  plyr::round_any(mean_prob_breaks[1:3], 0.01),
  max(rts_hex$mean_prob)
)

max_prob_breaks <- rgeoda::natural_breaks(4, rts_hex["max_prob"])
max_prob_breaks <- c(
  plyr::round_any(max_prob_breaks[1:3], 0.01),
  max(rts_hex$max_prob)
)

# breaks = tibble(
#   total_count_breaks = total_count_breaks,
#   total_area_breaks = total_area_breaks,
#   median_area_breaks = median_area_breaks,
#   mean_prob_breaks = mean_prob_breaks,
#   max_prob_breaks = max_prob_breaks
# )

# write_csv(
#   breaks,
#   "./visualization/data/mapping_class_breaks.csv"
# )
```

```{r}
rts_hex <- rts_hex |>
  mutate(
    total_count_class = classify_counts(total_count, total_count_breaks),
    total_area_class = classify_counts(total_area, total_area_breaks),
    median_area_class = classify_counts(median_area, median_area_breaks),
    mean_prob_class = classify_counts(mean_prob, mean_prob_breaks),
    max_prob_class = classify_counts(max_prob, max_prob_breaks),
    geometry_scaled_total_count = st_buffer(
      geometry,
      dist = calculate_buffer(total_count_class, 100000) * -1
    ), # geometry of scaled hexagons
    geometry_scaled_total_area = st_buffer(
      geometry,
      dist = calculate_buffer(total_area_class, 100000) * -1
    ), # geometry of scaled hexagons
    geometry_scaled_median_area = st_buffer(
      geometry,
      dist = calculate_buffer(median_area_class, 100000) * -1
    ), # geometry of scaled hexagons
    geometry_scaled_mean_prob = st_buffer(
      geometry,
      dist = calculate_buffer(mean_prob_class, 100000) * -1
    ), # geometry of scaled hexagons
    geometry_scaled_max_prob = st_buffer(
      geometry,
      dist = calculate_buffer(max_prob_class, 100000) * -1
    ) # geometry of scaled hexagons
  )

test_small_rts_hex <- test_small_rts_hex |>
  mutate(
    total_count_class = classify_counts(total_count, c(10, 100, 1000, 5000)),
    geometry_scaled_total_count = st_buffer(
      geometry,
      dist = calculate_buffer(total_count_class, 50000) * -1
    )
  )
```

```{r}
# geometry_cols <- rts_hex |>
#   select(starts_with("geometry")) |>
#   colnames()
# filepaths <- paste0(
#   "./visualization/data/rts_hex_2025_",
#   str_replace(geometry_cols, "geometry_scaled_", ""),
#   ".geojson"
# )
# filepaths[1] <- str_replace(filepaths[1], "_geometry", "")

# walk2(
#   geometry_cols,
#   filepaths,
#   ~ st_write(
#     rts_hex |>
#       st_set_geometry(.x),
#     .y,
#     delete_dsn = TRUE
#   )
# )
```

## Hex Inset (Several Options)

```{r}
inset1_hex <- rts_hex |>
  arrange(desc(total_count)) |>
  slice(1:6) |>
  mutate(id = 1:n(), .before = total_count)

inset1_bbox <- map(
  seq(1:6),
  ~ inset1_hex |>
    slice(.x) |>
    st_bbox()
)

inset1_figure_bbox <- map2(
  seq(1:6),
  inset1_bbox,
  ~ inset1_hex |>
    slice(.x) |>
    st_buffer(
      dist = (.y["xmax"] - .y["xmin"]) * 0.05,
      joinStyle = "MITRE",
      mitreLimit = (.y["xmax"] - .y["xmin"]) * 0.05
    ) |>
    st_bbox() |>
    st_as_sfc() |>
    st_as_sf() |>
    rename(geometry = x)
)

inset1_bbox <- map2_df(
  seq(1:6),
  inset1_bbox,
  ~ inset1_hex |>
    slice(.x) |>
    st_buffer(
      dist = (.y["xmax"] - .y["xmin"]) * 2,
      joinStyle = "MITRE",
      mitreLimit = (.y["xmax"] - .y["xmin"]) * 2
    ) |>
    st_bbox() |>
    st_as_sfc() |>
    st_as_sf() |>
    rename(geometry = x)
)

inset1_background <- map_df(
  seq(1:6),
  ~ inset1_bbox |>
    slice(.x) |>
    st_difference(inset1_hex |> slice(.x))
)

inset1_rts <- map_df(
  seq(1:6),
  ~ rts |>
    st_intersection(
      inset1_bbox |>
        slice(.x)
    )
) |>
  filter(row_number() == 1, .by = rts_id)

inset1_grid <- inset1_bbox |>
  st_make_grid(
    cellsize = 5000
  ) |>
  st_as_sf() |>
  rename(geometry = x) |>
  st_intersection(inset1_bbox |> st_union()) |>
  st_intersection(world_north |> st_union())

inset1_grid <- inset1_grid |>
  st_join(inset1_rts) |>
  summarize(
    total_count = n(),
    total_area = sum(area_m2),
    median_area = median(area_m2),
    mean_prob = sum(mean_prob * area_m2) / total_area,
    max_prob = sum(max_prob * area_m2) / total_area,
    .by = c(geometry)
  )

template <- inset1_grid |>
  vect() |>
  rast(res = 5000)

inset1_rast <- inset1_grid |>
  vect() |>
  rasterize(template, field = "total_count")
```

```{r}
# st_write(inset1_bbox, "./visualization/data/inset1_bbox.geojson")
# st_write(inset1_grid, "./visualization/data/inset1_grid.geojson")
# st_write(inset1_hex, "./visualization/data/inset1_hex.geojson")
# writeRaster(inset1_rast, "./visualization/data/inset1_grid_raster.tiff")
# st_write(inset1_rts, "./visualization/data/inset1_rts.geojson")
```

## Imagery Inset

Find the cluster of cells with the highest number of RTS using a neighborhood function.

```{r}
# Get bounding boxes of areas with a lot of RTS
inset2_bbox <- map_df(
  seq(1:nrow(inset1_hex)),
  ~ inset1_grid |>
    st_join(
      inset1_rast |>
        crop(
          inset1_hex |>
            slice(.x) |>
            vect()
        ) |>
        mask(
          inset1_hex |>
            slice(.x) |>
            vect()
        ) |>
        focal(w = 3, na.rm = TRUE, na.policy = "omit") |>
        as.data.frame(xy = TRUE) |>
        st_as_sf(coords = c("x", "y"), crs = st_crs(inset1_grid)) |>
        arrange(desc(focal_sum)) |>
        mutate(id = .x) |>
        slice_head(n = 1),
      left = FALSE
    ) |>
    st_buffer(
      dist = 5000,
      joinStyle = "MITRE",
      mitreLimit = 5000
    )
)

map(
  seq(1:nrow(inset1_hex)),
  ~ ggplot(
    # inset1_world
  ) +
    # geom_sf(
    #   color = 'gray85',
    #   fill = 'gray85'
    # ) +
    # geom_sf(data = inset1_rts) +
    geom_sf(
      data = inset1_grid |>
        st_crop(inset1_figure_bbox[[.x]]),
      aes(color = total_count, fill = total_count)
    ) +
    scale_color_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    scale_fill_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    new_scale_color() +
    geom_sf(
      data = world_ocean |>
        st_crop(inset1_figure_bbox[[.x]]),
      color = "transparent",
      fill = "#CCCCDD"
    ) +
    geom_sf(
      data = inset1_background |>
        slice(.x) |>
        st_crop(inset1_figure_bbox[[.x]]),
      color = "transparent",
      fill = "#FFFFFF",
      alpha = 0.75
    ) +
    geom_sf(
      data = inset1_hex |>
        slice(.x) |>
        st_crop(inset1_figure_bbox[[.x]]),
      fill = "transparent",
      linewidth = 0.5
    ) +
    geom_sf(
      data = inset2_bbox |>
        filter(id == .x),
      fill = "transparent",
      linewidth = 0.5
    ) +

    # Aesthetic Options
    coord_sf(expand = FALSE)
)

inset2_planet_grids <- map_df(
  seq(1:nrow(inset1_hex)),
  ~ planet_grids |>
    st_filter(
      inset2_bbox |>
        slice(.x),
      .predicate = st_intersects
    )
)

map(
  seq(1:nrow(inset1_hex)),
  ~ ggplot() +
    geom_sf(
      data = inset1_grid |>
        st_crop(inset2_bbox |> slice(.x)),
      aes(color = total_count, fill = total_count)
    ) +
    scale_color_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    scale_fill_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    geom_sf(
      data = inset1_ocean |>
        st_crop(inset2_bbox |> slice(.x)),
      color = "transparent",
      fill = "#CCCCDD"
    ) +
    geom_sf(
      data = inset2_planet_grids |>
        st_crop(inset2_bbox |> slice(.x)),
      fill = "transparent",
      linewidth = 0.5
    ) +
    geom_sf_text(
      data = inset2_planet_grids |>
        st_crop(inset2_bbox |> slice(.x)),
      aes(label = id)
    ) +
    geom_sf(
      data = inset1_rts |>
        st_crop(inset2_bbox |> slice(.x))
    ) +
    # geom_sf_text(data = inset1_planet_grids, aes(label = id)) +
    geom_sf(
      data = inset2_bbox |> slice(.x),
      fill = "transparent",
      linewidth = 1
    ) +

    # Aesthetic Options
    coord_sf(expand = FALSE)
)
```

```{r}
# st_write(
#   inset2_planet_grids,
#   "./visualization/data/inset2_planet_grids.geojson"
# )
```

## Region Hex

```{r}
region_hex <- hex_grid |>
  st_join(circumpolar_south, left = FALSE) |>
  summarise(
    geometry = st_union(geometry)
  )
```


```{r}
# st_write(
#   region_hex,
#   "./visualization/data/inference_region_hex_2025.geojson"
# )
```

# Map

## Visualization Prep

### Prepare Permafrost Legend

```{r}
#| fig-width: 4
#| fig-height: 1.3

ggplot() +
  geom_rect(
    aes(
      xmin = 0,
      xmax = 1,
      ymin = 0,
      ymax = 1
    ),
    color = "transparent",
    fill = "#6699CD"
  ) +
  geom_text(
    aes(
      x = 1.2,
      y = 0.5,
      label = "Permafrost"
    ),
    hjust = 0,
    size = 12
  ) +
  theme_void() +
  coord_fixed(
    clip = "off"
  ) +
  theme(
    plot.margin = margin(0.1, 2.6, 0.1, 0.1, unit = "in")
  )

# ggsave(
#   "./plots/circumpolar_maps/legends/permafrost.svg",
#   height = 1.3,
#   width = 4
# )
# ggsave(
#   "./plots/circumpolar_maps/legends/permafrost.png",
#   height = 1.3,
#   width = 4
# )
```

### Hex Legend Location

```{r}
size_legend_location <- c(
  left = 0,
  bottom = 0,
  right = 0.63,
  top = 0.12
)
```

## RTS Count Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
total_count_hexplot <- plot_hex_grid(
  rts_hex,
  geometry,
  geometry_scaled_total_count,
  total_count_class,
  rts_palettes[[2]]
)

total_count_legend <- create_hex_legend(
  rts_hex,
  "RTS Count",
  4,
  total_count_breaks,
  rts_palettes[[2]]
)

total_count_hexplot <- total_count_hexplot +
  inset_element(
    total_count_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
total_count_hexplot
```

```{r}
# ggsave(
#   "./plots/circumpolar_maps/legends/total_count_legend.png",
#   total_count_legend,
#   height = 1.5,
#   width = 5
# )

# ggsave(
#   "./plots/circumpolar_maps/legends/total_count_legend.svg",
#   total_count_legend,
#   height = 1.5,
#   width = 5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_total_count_map.png",
#   total_count_hexplot,
#   height = 6.5,
#   width = 6.5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_total_count_map.svg",
#   total_count_hexplot,
#   height = 6.5,
#   width = 6.5
# )
```

### Test 10,000 km^2 cell size rather than 100,000 km^2

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
test_count_hexplot <- plot_hex_grid(
  test_small_rts_hex,
  geometry,
  geometry_scaled_total_count,
  total_count_class,
  rts_palettes[[2]]
)

test_count_legend <- create_hex_legend(
  test_small_rts_hex,
  50000,
  "RTS Count",
  4,
  c(10, 100, 1000, 5000),
  rts_palettes[[2]]
)

test_count_hexplot <- test_count_hexplot +
  inset_element(
    test_count_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
test_count_hexplot
```

### Hex Inset

```{r}
map(
  seq(1:nrow(inset1_hex)),
  ~ ggplot(
    # inset1_world
  ) +
    # geom_sf(
    #   color = 'gray85',
    #   fill = 'gray85'
    # ) +
    # geom_sf(data = inset1_rts) +
    geom_sf(
      data = inset1_grid |>
        st_crop(inset1_figure_bbox[[.x]]),
      aes(color = total_count, fill = total_count)
    ) +
    scale_color_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    scale_fill_gradientn(
      name = "RTS Count",
      colors = rts_palettes[[1]]
    ) +
    geom_sf(
      data = world_ocean |>
        st_crop(inset1_figure_bbox[[.x]]),
      color = "transparent",
      fill = "#CCCCDD"
    ) +
    geom_sf(
      data = inset1_background |>
        slice(.x) |>
        st_crop(inset1_figure_bbox[[.x]]),
      color = "transparent",
      fill = "#FFFFFF",
      alpha = 0.75
    ) +
    geom_sf(
      data = inset1_hex |>
        slice(.x) |>
        st_crop(inset1_figure_bbox[[.x]]),
      fill = "transparent",
      linewidth = 0.5
    ) +

    # Aesthetic Options
    coord_sf(expand = FALSE)
)
```


### Imagery Inset

```{r}
ggplot() +
  geom_sf(data = inset2_world, color = "transparent") +
  geom_sf(data = inset2_ocean, color = "transparent", fill = "#CCCCDD") +
  geom_sf(data = inset2_rts) +

  # Aesthetic Options
  coord_sf(expand = FALSE)
```

## RTS Area Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
total_area_hexplot <- plot_hex_grid(
  rts_hex,
  geometry,
  geometry_scaled_total_area,
  total_area_class,
  rts_palettes[[2]]
)

total_area_legend <- create_hex_legend(
  rts_hex,
  "Total\nRTS Area",
  4,
  total_area_breaks,
  rts_palettes[[2]]
)

total_area_hexplot <- total_area_hexplot +
  inset_element(
    total_area_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
total_area_hexplot
```

```{r}
# ggsave(
#   "./plots/circumpolar_maps/legends/total_area_legend.png",
#   total_area_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/legends/total_area_legend.svg",
#   total_area_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_total_area_map.png",
#   total_area_hexplot,
#   height = 6.5,
#   width = 6.5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_total_area_map.svg",
#   total_area_hexplot,
#   height = 6.5,
#   width = 6.5
# )
```

## RTS Median Area Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
median_area_hexplot <- plot_hex_grid(
  rts_hex,
  geometry,
  geometry_scaled_median_area,
  median_area_class,
  rts_palettes[[2]]
)

median_area_legend <- create_hex_legend(
  rts_hex,
  "Median\nRTS Area",
  4,
  median_area_breaks,
  rts_palettes[[2]]
)

median_area_hexplot <- median_area_hexplot +
  inset_element(
    median_area_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
median_area_hexplot
```

```{r}
# ggsave(
#   "./plots/circumpolar_maps/legends/median_area_legend.png",
#   median_area_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/legends/median_area_legend.svg",
#   median_area_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_median_area_map.png",
#   median_area_hexplot,
#   height = 6.5,
#   width = 6.5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_median_area_map.svg",
#   median_area_hexplot,
#   height = 6.5,
#   width = 6.5
# )
```

## RTS Mean Probability Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
mean_prob_hexplot <- plot_hex_grid(
  rts_hex,
  geometry,
  geometry_scaled_mean_prob,
  mean_prob_class,
  rts_palettes[[2]]
)

mean_prob_breaks[[4]] <- ceiling(mean_prob_breaks[[4]] * 100) / 100


mean_prob_legend <- create_hex_legend(
  rts_hex,
  "Mean RTS\nProbability",
  4,
  mean_prob_breaks,
  rts_palettes[[2]]
)

mean_prob_hexplot <- mean_prob_hexplot +
  inset_element(
    mean_prob_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
mean_prob_hexplot
```

```{r}
# ggsave(
#   "./plots/circumpolar_maps/legends/mean_prob_legend.png",
#   mean_prob_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/legends/mean_prob_legend.svg",
#   mean_prob_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_mean_prob_map.png",
#   mean_prob_hexplot,
#   height = 6.5,
#   width = 6.5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_mean_prob_map.svg",
#   mean_prob_hexplot,
#   height = 6.5,
#   width = 6.5
# )
```

## RTS Max Probability Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
max_prob_hexplot <- plot_hex_grid(
  rts_hex,
  geometry,
  geometry_scaled_max_prob,
  max_prob_class,
  rts_palettes[[2]]
)

max_prob_breaks[[4]] <- ceiling(max_prob_breaks[[4]] * 100) / 100

max_prob_legend <- create_hex_legend(
  rts_hex,
  "Max RTS\nProbability",
  4,
  max_prob_breaks,
  rts_palettes[[2]]
)

max_prob_hexplot <- max_prob_hexplot +
  inset_element(
    max_prob_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
max_prob_hexplot
```

```{r}
# ggsave(
#   "./plots/circumpolar_maps/legends/max_prob_legend.png",
#   max_prob_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/legends/max_prob_legend.svg",
#   max_prob_legend,
#   height = 1.5,
#   width = 5.2
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_max_prob_map.png",
#   max_prob_hexplot,
#   height = 6.5,
#   width = 6.5
# )

# ggsave(
#   "./plots/circumpolar_maps/rts_max_prob_map.svg",
#   max_prob_hexplot,
#   height = 6.5,
#   width = 6.5
# )
```


## All Centroids

```{r}
#| echo: false
ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray90',
    fill = 'gray90'
  ) +
  # geom_sf(data = rts) +
  geom_sf(
    data = rts_pts,
    aes(color = conf_class),
    alpha = 0.2
  ) +
  scale_color_manual(
    name = "Confidence",
    labels = c("High", "Medium", "Low"),
    values = c("#FF0000", "#CC9999", "#CCCCCC")
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void()
```

## High Confidence Centroids Only

```{r}
#| echo: false
ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray90',
    fill = 'gray90'
  ) +
  # geom_sf(data = rts) +
  geom_sf(
    data = rts_pts |>
      filter(conf_class == "high"),
    color = red_max,
    alpha = 0.2
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void()
```


