---
title: "Training Data Distribution"
format: gfm
author: "Heidi Rodenhizer"
---

::: content-hidden
# Set Up

```{r}
#| include: false
library(googleCloudStorageR)
library(sf)
library(terra)
library(tidyterra)
library(arcgis)
library(tidyverse)
library(ggnewscale)
# library(biscale)
library(viridis)
library(patchwork)
```

```{r}
token <- auth_code()
```

```{r}
#| include: false
theme_set(theme_bw())
```


# Define Functions

```{r}
classify_counts <- function(count_data, breaks) {
  new_count_data <- case_when(
    count_data <= breaks[1] ~ 1,
    count_data <= breaks[2] ~ 2,
    count_data <= breaks[3] ~ 3,
    count_data <= breaks[4] ~ 4
  )
}
```

# Import Data

## Domain

```{r}
#| output: false
circumpolar <- st_read("./domain/circumpolar_domain.geojson") |>
  st_transform(6931)
circumpolar_south <- st_read("./domain/circumpolar_south_domain.geojson") |>
  st_transform(6931)
```

## Predictions

```{r}
# gs://rts-mapping-v2-usw1/inference/2025q3_south/products/south_rts_candidates.gpkg
rts <- st_read("../RTSMapping_v2_products/2025q3/south_rts_candidates.gpkg") |>
  st_transform(6931)
```

## Hex Grid

```{r}
hex_grid <- st_read("./visualization/data/training_hex_grid.geojson")
```

## Countries Map

```{r}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
crop_poly <- tibble(
  geometry = st_sfc(st_point(c(0, 90)), crs = 'EPSG:4326')
) %>%
  st_sf() %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  st_buffer(dist = 5250000) %>%
  smoothr::densify(n = 3) %>%
  st_transform(crs = 'EPSG:4326')

crop_bbox <- crop_poly |>
  st_transform(6931) |>
  st_bbox()

world_north <- world %>%
  st_make_valid() |>
  st_intersection(crop_poly) %>%
  st_transform(crs = st_crs(circumpolar)) %>%
  select(name_long, continent, region_un, subregion)

world_north$geom[4] <- st_buffer(world_north$geom[4], 0.25)

world_north <- world_north %>%
  st_union(by_feature = TRUE)

long_lines_from_coords <- function(lon, lat_min, lat_max) {
  line <- expand_grid(lon = lon, lat = c(lat_min, lat_max)) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lon)

  return(line)
}

long_lines <- seq(-180, 180, by = 60) %>%
  map(
    ~ long_lines_from_coords(.x, lat_min = st_bbox(crop_poly)[2], lat_max = 90)
  ) %>%
  bind_rows()

lat_lines_from_coords <- function(lat, lon_min, lon_max) {
  line <- expand_grid(lon = c(lon_min, lon_max), lat = lat) %>%
    st_as_sf(coords = c("lon", "lat"), crs = "EPSG:4326") %>%
    summarise(geometry = st_combine(geometry)) %>%
    st_cast("LINESTRING") %>%
    smoothr::densify(max_distance = 1) %>%
    mutate(label = lat)

  return(line)
}

lat_lines <- seq(50, 90, by = 10) %>%
  map(~ lat_lines_from_coords(.x, lon_min = -180, lon_max = 180)) %>%
  bind_rows()

world_ocean <- crop_bbox |>
  st_as_sfc() |>
  st_buffer(dist = 150000, joinStyle = "MITRE", mitreLimit = 150000) |>
  st_transform(st_crs(world_north)) |>
  st_difference(
    world_north |>
      st_union()
  )
```

## Cities

```{r}
cities <- st_read("../../../external_datasets/cities/cities.shp") |>
  st_transform(6931) |>
  filter(
    CITY_NAME %in%
      c("Anchorage", "Moscow", "Reykjavik", "Tromso", "Yakutsk", "Yellowknife")
  )
```

## Esri Multi-direction Hillshade

```{r}
hillshd_srvr <- arc_open(
  "https://elevation.arcgis.com/arcgis/rest/services/WorldElevation/Terrain/ImageServer",
  token = token
)

hillshd <- arc_raster(
  hillshd_srvr,
  xmin = -180,
  xmax = 180,
  ymin = 43.9,
  ymax = 90,
  crs = 4326,
  token = token
)

# hillshd = hillshd |>
#   crop(ext(vect(crop_poly)))
```

## Permafrost

```{r}
permafrost <- rast(
  "../rts_rep_analysis/input_data/NIEERPermafrost/NIEER_permafrost_dataset_released/NIEER_Probability.tif"
) |>
  aggregate(fact = 12) |>
  project("epsg:6931") |>
  mask(
    crop_poly |>
      st_transform(6931) |>
      vect()
  )
```

:::

# Prepare RTS Polygons for Mapping

## Calculate Centroids

```{r}
rts_pts <- rts |>
  st_centroid()

rts_pts_high <- rts_pts |>
  filter(conf_class == "high")
```

## Create Hex Grid

```{r}
rts_hex <- hex_grid |>
  st_join(rts_pts_high, left = FALSE) |>
  summarize(
    rts_count = n(),
    rts_area = sum(area_m2),
    mean_prob = sum(mean_prob * area_m2) / rts_area,
    .by = c(geometry)
  )
```

Take a look at distribution of counts and areas - consider a log scale for color ramps on these.

```{r}
ggplot(data = rts_hex) +
  geom_histogram(aes(x = rts_count)) +
  scale_x_continuous(trans = "log10")

ggplot(data = rts_hex) +
  geom_histogram(aes(x = rts_area)) +
  scale_x_continuous(trans = "log10")

quantile(rts_hex$rts_count)
nat_count_breaks <- rgeoda::natural_breaks(4, rts_hex["rts_count"])
count_breaks <- c(
  30,
  plyr::round_any(nat_count_breaks[1:2], 50),
  nat_count_breaks[3]
)

quantile(rts_hex$rts_area)
area_breaks <- rgeoda::natural_breaks(5, rts_hex["rts_area"])
```

```{r}
rts_hex <- rts_hex |>
  mutate(
    rts_count_class = classify_counts(rts_count, count_breaks),
    rts_area_class = classify_counts(rts_area, area_breaks),
    count_buffer = (
      sqrt((100000 * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
    ) -
      (
        sqrt(
          (100000 *
            ((rts_count_class - 1) / (max(rts_count_class) - 1) * 0.9 + 0.1) * # apply a log10 transformation
            2) /
            sqrt(3)
        ) /
          2 *
          1000 # radius of a circle of size (10,000 km^2 * count percentile) in m
      ), # the difference in m between a full size hex and the scaled hex (to use as a buffer in next line)
    geometry_count_scaled = st_buffer(geometry, dist = count_buffer * -1), # geometry of scaled hexagons
    area_buffer = (
      sqrt((100000 * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
    ) -
      (
        sqrt(
          (100000 *
            ((rts_area_class - 1) / (max(rts_area_class) - 1) * 0.9 + 0.1) * # apply a log10 transformation
            2) /
            sqrt(3)
        ) /
          2 *
          1000 # radius of a circle of size (10,000 km^2 * count percentile) in m
      ), # the difference in m between a full size hex and the scaled hex (to use as a buffer in next line)
    geometry_area_scaled = st_buffer(geometry, dist = area_buffer * -1) # geometry of scaled hexagons,
    # count_buffer = (
    #   sqrt((100000 * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
    # ) -
    #   (
    #     sqrt(
    #       (100000 *
    #         (log10(rts_count) / max(log10(rts_count)) * 0.98 + 0.02) * # apply a log10 transformation
    #         2) /
    #         sqrt(3)
    #     ) /
    #       2 *
    #       1000 # radius of a circle of size (10,000 km^2 * count percentile) in m
    #   ), # the difference in m between a full size hex and the scaled hex (to use as a buffer in next line)
    # geometry_count_scaled = st_buffer(geometry, dist = count_buffer * -1), # geometry of scaled hexagons
    # area_buffer = (
    #   sqrt((100000 * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
    # ) -
    #   (
    #     sqrt(
    #       (100000 *
    #         (log10(rts_area) / max(log10(rts_area)) * 0.98 + 0.02) * # apply a log10 transformation
    #         2) /
    #         sqrt(3)
    #     ) /
    #       2 *
    #       1000 # radius of a circle of size (10,000 km^2 * count percentile) in m
    #   ), # the difference in m between a full size hex and the scaled hex (to use as a buffer in next line)
    # geometry_area_scaled = st_buffer(geometry, dist = area_buffer * -1) # geometry of scaled hexagons
  )
```

## Region Hex

```{r}
region_hex <- hex_grid |>
  st_join(circumpolar_south, left = FALSE) |>
  summarise(
    geometry = st_union(geometry)
  )
```

# Prep First Zoom Inset

```{r}
example_hex <- rts_hex |>
  filter(rts_count == max(rts_count))

example_rts <- rts |>
  st_intersection(example_hex)

example_bbox <- st_bbox(example_hex)

example_world <- world_north |>
  st_crop(
    example_hex |>
      st_buffer(
        dist = (example_bbox["xmax"] - example_bbox["xmin"]) * 0.05,
        joinStyle = "MITRE",
        mitreLimit = (example_bbox["xmax"] - example_bbox["xmin"]) * 0.05
      )
  )
```

# Map

## Prepare Elevation/Permafrost Blend

```{r}
blend_overlay <- function(a, b) {
  case_when(
    a < 0.5 ~ 2 * a * b,
    is.na(b) ~ a,
    TRUE ~ (1 - 2 * (1 - a) * (1 - b))
  )
}

elev_perm_overlay <- c(hillshd, project(permafrost, hillshd))
names(elev_perm_overlay) <- c("hillshd", "perm_prob")
elev_perm_overlay["blend"] <- lapp(elev_perm_overlay[[1:2]], blend_overlay)
```

## Colors

```{r}
reds <- colorRampPalette(c("#FFFFFF", "#AE3A4E"))
reds_10 <- reds(10)
red_min <- reds_10[2]
red_max <- reds_10[10]
```

## All Centroids

```{r}
#| echo: false
ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray90',
    fill = 'gray90'
  ) +
  # geom_sf(data = rts) +
  geom_sf(
    data = rts_pts,
    aes(color = conf_class),
    alpha = 0.2
  ) +
  scale_color_manual(
    name = "Confidence",
    labels = c("High", "Medium", "Low"),
    values = c("#FF0000", "#CC9999", "#CCCCCC")
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void()
```

## High Confidence Centroids Only

```{r}
#| echo: false
ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray90',
    fill = 'gray90'
  ) +
  # geom_sf(data = rts) +
  geom_sf(
    data = rts_pts |>
      filter(conf_class == "high"),
    color = red_max,
    alpha = 0.2
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void()
```

## Circumpolar Hex

### Size Legend

```{r}
hex_legend_data <- rts_hex |>
  slice(rep(1, 4))
bounds <- st_bbox(hex_legend_data$geometry)
hex_legend_data <- hex_legend_data |>
  mutate(
    geometry = geometry - c(bounds$xmin, bounds$ymin),
    count = c(
      4,
      3,
      2,
      1
    ),
    buffer = (
      sqrt((100000 * 2) / sqrt(3)) / 2 * 1000 # radius of a circle of size 10,000 km^2 in m
    ) -
      (
        sqrt(
          (100000 *
            ((count - 1) / (max(count) - 1) * 0.9 + 0.1) *
            2) /
            sqrt(3)
        ) /
          2 *
          1000 # radius of a circle of size (10,000 km^2 * count percentile) in m
      ), # the difference in m between a full size hex and the scaled hex (to use as a buffer in next line)
    geometry_scaled = st_buffer(geometry, dist = buffer * -1) # geometry of scaled hexagons
  ) |>
  rowwise() |>
  mutate(
    geometry_scaled = geometry_scaled - c(0, st_bbox(geometry_scaled)$ymin)
  ) |>
  select(count, buffer, geometry, geometry_scaled)

arrow <- tibble(
  x = 0 - 100000,
  y = 0 + 15000,
  xend = 0 - 100000,
  yend = st_bbox(hex_legend_data)["ymax"] - 100000
)

labels <- tibble(
  x = rep(-100000, 4),
  xend = c(
    st_bbox(hex_legend_data)["xmax"] -
      (st_bbox(hex_legend_data)["xmax"] - st_bbox(hex_legend_data)["xmin"]) / 2,
    NA,
    NA,
    st_bbox(hex_legend_data)["xmax"] -
      (st_bbox(hex_legend_data)["xmax"] - st_bbox(hex_legend_data)["xmin"]) / 2
  ),
  y = seq(
    st_bbox(hex_legend_data$geometry_scaled[[4]])["ymax"],
    st_bbox(hex_legend_data)["ymax"],
    length.out = 4
  ),
  breaks = c(count_breaks),
  label = paste0("<=", count_breaks)
)

size_legend <- ggplot() +
  geom_sf(
    data = hex_legend_data,
    aes(
      geometry = geometry_scaled,
      fill = count
    ),
    color = "transparent"
  ) +
  scale_fill_gradient(
    low = red_min,
    high = red_max,
    trans = "log10"
  ) +
  # geom_segment(
  #   data = arrow,
  #   aes(
  #     x = x,
  #     y = y,
  #     xend = xend,
  #     yend = yend
  #   ),
  #   arrow = arrow(length = unit(3, "points"), angle = 45),
  #   linewidth = 0.2
  # ) +
  geom_segment(
    data = labels,
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = y
    ),
    linewidth = 0.2
  ) +
  geom_text(
    data = labels,
    aes(
      x = x,
      y = y,
      label = label
    ),
    hjust = 1,
    nudge_x = -50000,
    size = 1.75
  ) +
  geom_text(
    aes(
      x = st_bbox(hex_legend_data)["xmax"] -
        (st_bbox(hex_legend_data)["xmax"] - st_bbox(hex_legend_data)["xmin"]) /
          2,
      y = max(labels$y) + 100000,
      label = "Total Count"
    ),
    hjust = 0.5,
    size = 1.75,
    # angle = 90
  ) +
  coord_sf(clip = "off") +
  theme_void() +
  theme(
    legend.position = "none"
  )
# size_legend
```

### RTS Count Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
rts_count_hexplot <- ggplot(world_north) +
  # Background (ocean, lon, lat, land)
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray85',
    fill = 'gray85'
  ) +
  geom_spatraster(data = hillshd) +
  scale_fill_gradient(
    low = "#CCCCCC",
    high = "#FFFFFF",
    na.value = "transparent"
  ) +
  new_scale_fill() +
  geom_sf(
    data = world_ocean,
    fill = "white",
    color = "transparent"
  ) +

  # Permafrost
  geom_spatraster(
    data = permafrost,
    aes(fill = NIEER_Probability),
    # maxcell = 10e+06,
    alpha = 0.75
  ) +
  scale_fill_gradient(
    low = "#d8ebfd",
    high = "#88BBEE",
    na.value = "transparent"
  ) +
  new_scale_fill() +

  # Cities
  geom_sf(
    data = cities,
    shape = 21,
    color = "black",
    fill = "white",
    size = 2
  ) +
  geom_sf(
    data = cities,
    color = "black",
    size = 0.5
  ) +
  geom_sf_text(
    data = cities,
    aes(label = CITY_NAME),
    hjust = 1,
    position = position_nudge(x = -150000),
    color = "black",
    size = 2
  ) +

  # Country borders
  geom_sf(
    color = 'white',
    fill = 'transparent',
    linewidth = 0.1
  ) +

  # RTS Hex
  geom_sf(
    data = rts_hex,
    aes(geometry = geometry_count_scaled, fill = rts_count_class),
    color = "gray10",
    alpha = 0.8
  ) +
  scale_fill_gradient(
    name = "RTS Count",
    low = red_min,
    high = red_max,
    trans = "log10"
  ) +

  # # RTS Hex Outlines
  # geom_sf(
  #   data = rts_hex,
  #   aes(geometry = geometry),
  #   fill = "transparent",
  #   color = "gray95",
  #   linewidth = 0.3
  # ) +

  # Inference Region
  geom_sf(
    data = region_hex,
    color = "black",
    fill = "transparent",
    linewidth = 0.3
  ) +

  # Manual Circular Plot Border
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +

  # Aesthetic Options
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf(
    xlim = c(crop_bbox["xmin"], crop_bbox["xmax"]),
    ylim = c(crop_bbox["ymin"], crop_bbox["ymax"])
  ) +
  theme_void() +
  theme(
    # legend.text = element_text(size = 5, margin = margin(0, 3, 0, 0)),
    # legend.text.position = "left",
    # legend.position = "inside",
    # legend.position.inside = c(1, 0.17),
    # legend.justification = c(1, 0),
    # legend.key.size = unit(13, "points"),
    # legend.key.spacing.x = unit(0, "points")
    legend.position = "none"
  )

size_legend_location <- c(
  left = 0.75,
  bottom = 0.1,
  right = 0.85,
  top = 0.21
)

rts_count_hexplot <- rts_count_hexplot +
  inset_element(
    size_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
rts_count_hexplot
```

```{r}
ggsave(
  "./plots/rts_count_map.png",
  rts_count_hexplot,
  height = 6.5,
  width = 6.5
)

ggsave(
  "./plots/rts_count_map.svg",
  rts_count_hexplot,
  height = 6.5,
  width = 6.5
)
```

## 1 Hex Zoom

```{r}
ggplot(
 example_world
) +
  geom_sf(
    color = 'gray85',
    fill = 'gray85'
  ) +
  geom_sf(data = example_rts) +
  geom_sf(data = example_hex, fill = "transparent", linewidth = 0.5) +

  # Aesthetic Options
  scale_x_continuous(expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
```



### RTS Area Map

```{r}
#| fig-width: 6.5
#| fig-height: 6.5
rts_count_hexplot <- ggplot(world_north) +
  geom_sf(
    data = long_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    data = lat_lines,
    color = 'gray85',
    linewidth = 0.1
  ) +
  geom_sf(
    color = 'gray90',
    fill = 'gray90'
  ) +
  geom_sf(
    data = rts_hex,
    aes(geometry = geometry_count_scaled, fill = rts_area),
    color = "transparent",
    alpha = 1
  ) +
  geom_sf(
    data = rts_hex,
    aes(geometry = geometry),
    fill = "transparent",
    color = "gray95",
    linewidth = 0.5
  ) +
  scale_fill_gradient(
    name = "RTS Area",
    low = red_min,
    high = red_max,
    trans = "log10"
  ) +
  geom_sf(
    data = crop_poly,
    color = 'black',
    fill = 'transparent',
    linewidth = 0.25
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  coord_sf() +
  theme_void() +
  theme(
    legend.text = element_text(size = 10, margin = margin(0, 3, 0, 0)),
    legend.text.position = "left",
    legend.position = "inside",
    legend.position.inside = c(1, 0.17),
    legend.justification = c(1, 0),
    legend.key.size = unit(13, "points"),
    legend.key.spacing.x = unit(0, "points")
  )

size_legend_location <- c(
  left = 0.75,
  bottom = 0.019,
  right = 0.8,
  top = 0.074
)

rts_count_hexplot <- rts_count_hexplot +
  inset_element(
    size_legend,
    left = size_legend_location["left"],
    bottom = size_legend_location["bottom"],
    right = size_legend_location["right"],
    top = size_legend_location["top"]
  )
rts_count_hexplot
```

```{r}
ggsave(
  "./plots/rts_area_map.png",
  rts_area_hexplot,
  height = 6.5,
  width = 6.5
)

ggsave(
  "./plots/rts_area_map.svg",
  rts_area_hexplot,
  height = 6.5,
  width = 6.5
)
```

